Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-07-10 prop_letter.tex
After meeting...

\title{Interval-Valued Quantum Probabilities: Contextuality and the Born Rule}

@@

\begin{abstract}~
\begin{description}
\item [{Background}] In a world of bounded resource, any quantum measurement
can be only performed in finite precision.
\item [{Method}] We model the idea of the finite precision measurement
as finite number of intervals, and replace quantum real-valued probability
measures by interval-valued ones.
\item [{Results}] Quantum interval-valued probability measures provide
a unified framework to interpret Gleason's theorem and the Kochen-Specker
theorem as the limiting cases. For the cases between them, we found
some legitimate quantum interval-valued probability measures which
cannot be explained by the Born rule. While restricting on only commuting
projectors, the interval-valued probability measures become classical,
and these non-Born interval-valued probability measures disappear.
\item [{Conclusions}] Our research provides a new insight of finite precision
measurement and non-commutativity when formulating quantum theory.
Also, the Born rule may not be as nature as we used to believe in
the world of finite resource.
\end{description}
\end{abstract}

@@

In this letter, we extend our previous work on the foundations of
computability in quantum mechanics by exploring the application of
interval-valued probability measures (IVPM) to achieving a coherent
formulation of finite-resource quantum mechanics.

There are long-standing debates in the foundations of quantum
mechanics regarding the tension between finite precision measurement
and contextuality\nb\cite{BarrettKent2004,Appleby_2005}.  For example,
Meyer\nb\cite{PhysRevLett.83.3751} and Mermin\nb\cite{Mermin1999} disagree
on whether finite-precision measurements invalidate the Kochen-Specker
theorem\nb\cite{kochenspecker1967,peres1995quantum}, and discussion
continues on whether contextuality is experimentally
verifiable\nb\cite{Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}. The
fundamental question is how to make mathematical theories of quantum
mechanics that are intrinsically, rather than only implicitly,
consistent with the resources available for the realistic accuracy of
an actual measurement.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fuzzy Measurements}

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections of
uncertain events\nb\cite{Kolmogorov1950}. In the quantum case, the
events of interest are specified by \emph{projection operators}
$P$. These include the empty projection\nb$\mathbb{0}$, the identity
projection\nb$\mathbb{1}$, projections of the form $\proj{\psi}$ where
$\ket{\psi}$ is a pure quantum state (an element of a Hilbert space),
sums of \emph{orthogonal} projections $P_0$ and $P_1$ such
$P_0P_1=\mathbb{0}$, and products of \emph{commuting} projections
$P_0$ and $P_1$ such $P_0P_1=P_1P_0$. In a well-defined quantum
probability
space\nb\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010},
each event $P_{i}$ is mapped using a probability measure
$\mu : P \rightarrow [0,1]$ to a probability $\mu(P_{i})$ subject to
the following constraints: $\mu(\mathbb{0})=0$, $\mu(\mathbb{1})=1$,
$\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$, and for each pair
of \emph{orthogonal} projections $P_{0}$ and $P_{1}$, we have:
\begin{equation}
{\mu}\left(P_{0}+P_{1}\right)={\mu}\left(P_{0}\right)+{\mu}\left(P_{1}\right)\,.\label{eq:QuantumProbability-Addition}
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Intervals of Uncertainty}

We will start by reviewing IVPMs in the classical setting and then
propose our quantum generalization. In the classical setting, there
are several proposals for ``imprecise
probabilities''~\citep{Dempster1967,Shafer1976,GilboaSchmeidler1994,Weichselberger2000,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
Although these proposals differ in some details, they all share the
fact that the probability $\mu(E)$ of an event~$E$ is generalized from
a single real number to an interval of certainty $[\ell,r]$
where\nb$\ell$ intuitively corresponds to the strength of evidence for
the event and~$r$ corresponds to the strength of evidence against the
event. The collection of intervals used in the definition of a
probability measure must satisfy certain properties which we explain
next.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Kochen-Specker}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Gleason}

In the classical setting, measurements done with limited resources and
poor precision may not uniquely identify the true state of a
system. However, there is always at least one state that is consistent
with the imprecise
measurements~\cite{Shapley1971,GilboaSchmeidler1994,Grabisch2016}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

Foundational concepts in quantum mechanics, such as Gleason and
Kochen-Specker theorems, rely in a very subtle way on the use of
unbounded resources. By assuming infinitely precise measurements,
these two insightful theorems underline two fundamental aspects of
quantum mechanics. On the one hand, Gleason's result establishes a
relationship between quantum states and probabilities uniquely defined
by Born's rule, on the other, Kochen-Specker unveils a distinctive
aspect of physical reality, the fact that it is contextual.  Our goal
in this paper has been to analyze the physical consequences of a
mathematical framework that allows for finite precision measurements
by introducing the concept of quantum interval-valued probability. The
latter incorporates uncertainty in the measurement results by defining
fuzzy probability measures, and includes standard quantum measurement
as the particular instance of sharp, infinitely precise, intervals.
In addition, we showed how these two theorems emerge as limiting cases
of this very same framework, thus connecting two seemingly unrelated
aspects of quantum physics. Most importantly, we concluded that
arbitrary finite precision measurement conditions nullify the main
tenets of both theorems.

>>>>

\title{Quantum Interval-Valued Probability: Contextuality and the Born Rule}

@@

\begin{abstract}~
\begin{description}
\item [{Background}] In a world of bounded resource, any quantum measurement
can be only performed in finite precision.
\item [{Method}] We model the idea of the finite precision measurement
as finite number of intervals, and replace quantum real-valued probability
measures by interval-valued ones.
\item [{Results}] Quantum interval-valued probability measures provide
a unified framework to interpret Gleason's theorem and the Kochen-Specker
theorem as the limiting cases. For the cases between them, we found
some legitimate quantum interval-valued probability measures which
cannot be explained by the Born rule. While restricting on only commuting
projectors, the interval-valued probability measures become classical,
and these non-Born interval-valued probability measures disappear.
\item [{Conclusions}] Our research provides a new insight of finite precision
measurement and non-commutativity when formulating quantum theory.
Also, the Born rule may not be as nature as we used to believe in
the world of finite resource.

\end{description}
\end{abstract}

@@

In this paper, we extend our previous work on the foundations of
computability in quantum mechanics by exploring the application of
interval-valued probability measures (IVPM) to achieving a coherent
formulation of finite-resource quantum mechanics.  In particular, we quantify the concept of imperfect quantum measurement and its impact on the interpretation of the Kochen-Specker and Gleason theorems and their implications for the foundations of quantum mechanics.

There are long-standing debates in the foundations of quantum
mechanics regarding the tension between finite precision measurement
and contextuality\nb\cite{BarrettKent2004,Appleby_2005}.  For example,
Meyer\nb\cite{PhysRevLett.83.3751} and Mermin\nb\cite{Mermin1999} disagree
on whether finite-precision measurements invalidate the Kochen-Specker
theorem\nb\cite{kochenspecker1967,peres1995quantum}, and discussion
has continued on whether contextuality is consistent with  experimental outcomes\nb\cite{Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}. The
fundamental question is how to make mathematical theories of quantum mechanics that are intrinsically, rather than only implicitly, consistent with the resources available for the realistic accuracy of an actual measurement.

Here we develop a mathematical framework that includes the uncertainties of finite precision and imperfections in the quantum measurement process.  Our first observation is that essential objectives of our program to
achieve a computable theory of quantum mechanics and finite-resource
measurement, attempted previously with computable number
systems\nb\cite{usat,geometry2013apsrev4,DQT2014}, may be achievable by
extending classical interval-valued probability
measures\nb\cite{JamisonLodwick2004} to the quantum domain.  We begin by axiomatizing a quantum interval-valued probability measure (QIVPM) framework, which will provide a quantitative description of the impact of finite precision on the foundations of quantum mechanics.

% exploiting the fact that a density matrix approach can be
% considered as equivalent to a description of quantum states
% using probability  measures\nb\cite{10.2307/2308516,Varadarajan2008}. %Mackey=10.2307

In particular, the effect of finite precision on contextuality, and the Kochen-Specker theorem, is shown to have a resolution in terms of the quantum interval valued probability measure framework that we introduce.  To proceed, we use a formulation of the Mermin square that uses the
expectation value of an observable with respect
to a QIVPM can be defined to be consistent with the constraints on the
expectation value of the same observable expected from continuous
quantum mechanics.  We then examine Gleason's
theorem\nb\cite{gleason1957,Redhead1987-REDINA,peres1995quantum} and the
Kochen-Specker theorem in the context of QIVPM's, and conclude that
they correspond to the limit of an infinite number of precise
intervals, and the set of cryptodeterministic intervals (the set\nb$\{\imposs,\necess\}$),
respectively.

\begin{itemize}
\item Develop a framework that includes the uncertainties of finite precision in the quantum measurement process.
\item Provide a quantitative description of the impact of finite precision on the foundations of quantum mechanics.
\item In particular, the effect of finite precision on contextuality is shown to have a resolution in terms of the quantum interval valued probability measure framework that we introduce.
\item A key aspect of quantum measurement is its stochastic nature as reflected in Born's rule.  Gleason's theorem establishes the existence of a unique state (in Hilbert spaces of dimension $\ge 3$), the probabilities of whose outcomes are computed from Born's rule assuming infinite precision.  Then it is legitimate to question what happens when infinite precision cannot be achieved.
\item 
\end{itemize}

@@

The paper is organized as follows.   In Sec.\ \ref{sec:fuzzy} we begin by introducing the foundations of quantum probability space and fuzzy measurement.  We then move on to introduce the concept of quantum interval valued probability in Sec.\ \ref{sec:Interval-Uncertainty}.  This QIVPM will provide the mathematical framework to quantify the impact of finite precision measurements on quantum mechanics.  For instance, in Sec.\ \ref{sec:Kochen-Specker},  we quantify the domain of validity of a  contextual measurement, thus addressing the conditions under which the Kochen-Specker theorem applies.  This provides a way of resolving the Meyer-Mermin debate, which is the main result of our Theorem \ref{cor:Kochen-Specker-IVPM}.  In Sec.\ \ref{sec:Gleason}, we study Gleason's theorem in the context of imprecise measurements, concluding that QIVPM provides a framework with quantitative bounds in which Gleason's theorem, while formally invalid, holds asymptotically.  Finally, we present our conclusions in Sec.~\ref{sec:Conclusion}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fuzzy Measurements}
 \label{sec:fuzzy}
A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections of
uncertain events\nb\cite{Kolmogorov1950}. In the quantum case, the
events of interest are specified by \emph{projection operators}
$P$. These include the empty projection\nb$\mathbb{0}$, the identity
projection\nb$\mathbb{1}$, projections of the form $\proj{\psi}$ where
$\ket{\psi}$ is a pure quantum state (an element of a Hilbert space),
sums of \emph{orthogonal} projections $P_0$ and $P_1$ such
$P_0P_1=\mathbb{0}$, and products of \emph{commuting} projections
$P_0$ and $P_1$ such $P_0P_1=P_1P_0$. In a well-defined quantum
probability
space\nb\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010},
each event $P_{i}$ is mapped using a probability measure
$\mu : P \rightarrow [0,1]$ to a probability $\mu(P_{i})$ subject to
the following constraints: $\mu(\mathbb{0})=0$, $\mu(\mathbb{1})=1$,
$\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$, and for each pair
of \emph{orthogonal} projections $P_{0}$ and $P_{1}$, we have:
\begin{equation}
{\mu}\left(P_{0}+P_{1}\right)={\mu}\left(P_{0}\right)+{\mu}\left(P_{1}\right)\,.\label{eq:QuantumProbability-Addition}
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Intervals of Uncertainty}
  \label{sec:Interval-Uncertainty}
We will start by reviewing IVPMs in the classical setting and then
propose our quantum generalization. In the classical setting, there
are several proposals for ``imprecise
probabilities''~\citep{Dempster1967,Shafer1976,GilboaSchmeidler1994,Weichselberger2000,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
Although these proposals differ in some details, they all share the
fact that the probability $\mu(E)$ of an event~$E$ is generalized from
a single real number to an interval of certainty $[\ell,r]$
where\nb$\ell$ intuitively corresponds to the strength of evidence for
the event and~$r$ corresponds to the strength of evidence against the
event. The collection of intervals used in the definition of a
probability measure must satisfy certain properties which we explain
next.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Kochen-Specker} 
  \label{sec:Kochen-Specker}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Gleason}
  \label{sec:Gleason}
In the classical setting, measurements done with limited resources and
poor precision may not uniquely identify the true state of a
system. However, there is always at least one state that is consistent
with the imprecise
measurements~\cite{Shapley1971,GilboaSchmeidler1994,Grabisch2016}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
  \label{sec:Conclusion}

Foundational concepts in quantum mechanics, such as the Kochen-Specker and
Gleason theorems, rely in a very subtle way on the use of
unbounded resources. By assuming infinitely precise measurements,
these two insightful theorems underlie two fundamental aspects of
quantum mechanics. On the one hand, Kochen-Specker's result reveals a distinctive aspect of physical reality, the fact that it is contextual, and on the other hand, Gleason's theorem establishes a
relationship between quantum states and probabilities uniquely defined
by Born's rule.  Our goal
in this paper has been to analyze the physical consequences of a
mathematical framework that allows for finite precision measurements
by introducing the concept of quantum interval-valued probability. This
framework incorporates uncertainty in the measurement results by defining
fuzzy probability measures, and includes standard quantum measurement
as the particular instance of sharp, infinitely precise, intervals.
In addition, we showed how these two theorems emerge as limiting cases
of this same framework, thus connecting two seemingly unrelated
aspects of quantum physics. We noted that
{\it arbitrary\/} finite precision measurement conditions can nullify the main
tenets of both theorems. However, by carefully specifying experimental uncertainties, we were able to establish rigorous bounds on the validity of these two theorems. Therefore, we have established a context in which infinite precision quantum mechanical theories can be reconciled with finite precision quantum mechanical measurements, and have provided a possible resolution of the Meyer-Mermin debate on the impact of finite precision on the Kochen-Specker theorem \cite{PhysRevLett.83.3751,Mermin1999}.
